---
title: "updated models"
author: "Tony Carilli"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      options(digits = 3, scipen = 999))
library(tidyverse)
library(plm)
library(bibtex)
library(stargazer)
df <- # data
  here::here("data", "covid_data_final.csv") %>% 
  read_csv()
```

# Table 2 Regressions

```{r table-2, eval = FALSE}


df %>%
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  )


df %>%
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )


df %>%
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )


```

# Table 3 Regressions

```{r table-3, results='asis', message=FALSE, eval = FALSE}
model_penn_state <- 
  df %>%
  mutate(sc = scp_version_of_penn_state_index) %>% 
  plm(
    cases_new ~ stringency_index_lag * sc +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  )

model_putnam <- 
  df %>%
  mutate(sc = putnam_index) %>% 
  plm(
    cases_new ~ stringency_index_lag * sc +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  )

model_alesina <- 
  df %>%
  mutate(sc = alesina_la_ferrara_social_capital_group) %>% 
  plm(
    cases_new ~ stringency_index_lag * sc +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  )

stargazer(
  model_alesina,
  model_putnam,
  model_penn_state,
  type = "html",
  df = FALSE,
  header = FALSE,
  covariate.labels = c(
    "Stringency Index",
    "Stringency Index squared",
    "Stringency Index * Social Capital Index"
  ),
  title = "Alternate Social Capital Measures",
  column.labels = c(
    "Alesina La Ferrara",
    "Putnam Index",
    "Penn State Index"
  ),
  dep.var.labels = "Daily New Cases",
  notes = "Each model estimtate with two-way fixed effects."
)
```

# Table 4 Regressions

```{r table-4, results='asis', message=FALSE, eval = FALSE}
model_penn_state_one_way <-
  df %>%
  mutate(sc = scp_version_of_penn_state_index) %>%
  plm(
    cases_new ~ stringency_index_lag * sc +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )

model_putnam_one_way <-
  df %>%
  mutate(sc = putnam_index) %>%
  plm(
    cases_new ~ stringency_index_lag * sc +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )

model_alesina_one_way <-
  df %>%
  mutate(sc = alesina_la_ferrara_social_capital_group) %>%
  plm(
    cases_new ~ stringency_index_lag * sc +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )

stargazer(
  model_alesina_one_way,
  model_putnam_one_way,
  model_penn_state_one_way,
  type = "html",
  header = FALSE,
  df = FALSE,
  # omit = c(
  #   "median_household_income",
  #   "percent_adults_with_ba",
  #   "percent_in_fair_or_poor_health",
  #   "percent_black",
  #   "density",
  #   "percent_rural",
  #   "percent_65",
  #   "average_temperature",
  #   "governor_political_affiliation"
  # ),
  covariate.labels = c(
    "Stringency Index",
    "Stringency Index squared",
    "Social Capital Index",
    "Stringency Index * Social Capital Index"
  ),
  title = "Alternate Social Capital Measures",
  column.labels = c(
    "Alesina La Ferrara",
    "Putnam Index",
    "Penn State Index"
  ),
  order = c(1, 3, 2, 4),
  dep.var.labels = "Daily New Cases per Million",
  notes = "One-way fixed effects with controls not reported."
)
```

# Table 5 Regressions

```{r table-5, results='asis', message=FALSE, eval = FALSE}
model_oneway_with_controls_pre_july_4 <- 
  df %>% 
  filter(date <= "2020-07-04") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )
  
  model_oneway_with_controls_post_july_4 <- 
  df %>% 
  filter(date > "2020-07-04") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  )
  
stargazer(
  model_oneway_with_controls_pre_july_4,
  model_oneway_with_controls_post_july_4,
  type = "html",
  header = FALSE,
  df = FALSE,
  # omit = c(
  #   "median_household_income",
  #   "percent_adults_with_ba",
  #   "percent_in_fair_or_poor_health",
  #   "percent_black",
  #   "density",
  #   "percent_rural",
  #   "percent_65",
  #   "average_temperature",
  #   "governor_political_affiliation"
  # ),
  covariate.labels = c(
    "Stringency Index",
    "Stringency Index squared",
    "Social Capital Index",
    "Stringency Index * Social Capital Index"
  ),
  order = c(1, 3, 2, 4),
  title = "Regression Results pre and post July 4",
  column.labels = c("Mar - July 4",
                    "July 5 - "),
  dep.var.labels = "Daily New Cases per Million",
  notes = "One-way fixed effects with controls not reported."
) 
```

# Quarterly Regressions - State-level index - one way

```{r eval=false}
df %>% 
  filter(date < "2020-06-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>% 
  filter(date >= "2020-06-01", date < "2020-09-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>% 
  filter(date >= "2020-09-01", date < "2020-12-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>% 
  filter(date >= "2020-12-01", date < "2021-03-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2) +
      median_household_income +
      percent_adults_with_ba +
      percent_in_fair_or_poor_health +
      percent_black +
      density +
      percent_rural +
      percent_65 +
      average_temperature +
      governor_political_affiliation,
    data = .,
    effect = "time",
    index = "state",
    model = "within"
  ) %>% 
  summary()
```

# Quarterly Regressions - State-level index - two ways

```{r eval=FALSE}
df %>%
  filter(date < "2020-06-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  filter(date >= "2020-06-01", date < "2020-09-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  filter(date >= "2020-09-01", date < "2020-12-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  filter(date >= "2020-12-01", date < "2021-03-01") %>% 
  plm(
    cases_new ~ stringency_index_lag * state_level_index +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()
```

# Sub-indices regressions

```{r eval=FALSE}
df %>%
  plm(
    cases_new ~ stringency_index_lag * family_unity +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * family_interaction +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * social_support +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * community_health +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * institutional_health +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * philanthropic_health +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * average_number_of_close_friends +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_who_trust_all_most_neighbors +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * membership_organizations_p_1_000 +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * non_religious_non_profits_plus_religious_congregations_p_1_000 +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_participated_in_demonstration +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_who_volunteered_for_a_group +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_attended_public_meeting +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_worked_with_neighbors_to_improve_fix_something +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_served_on_committee_or_as_group_officer +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()

df %>%
  plm(
    cases_new ~ stringency_index_lag * percent_have_some_great_confidence_in_media +
      I(stringency_index_lag ^ 2),
    data = .,
    effect = "twoways",
    index = "state",
    model = "within"
  ) %>% 
  summary()
```

# Index and sub-index regressions complete two-way

```{r}
sc_names <-
  df %>%
  names() 

soc_cap_measures <- sc_names[c(20:23, 52:70)]  
  
```

```{r}

models_by_social_capital_measure <- list()

for (i in seq_along(soc_cap_measures)) {
  # for(i in 1:1)  {
  # sc_i <- soc_cap_measures[i]
  
  models_by_social_capital_measure[[i]] <-
    df %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2),
      data = .,
      effect = "twoways",
      index = "state",
      model = "within"
    ) %>%
    summary()
}

names(models_by_social_capital_measure) <- soc_cap_measures

```

# Index and sub-index regressions complete one-way

```{r}

models_by_social_capital_measure_one_way <- list()

for (i in seq_along(soc_cap_measures)) {
  models_by_social_capital_measure_one_way[[i]] <-
    df %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2) +
        median_household_income +
        percent_adults_with_ba +
        percent_in_fair_or_poor_health +
        percent_black +
        density +
        percent_rural +
        percent_65 +
        average_temperature +
        governor_political_affiliation,
      data = .,
      effect = "time",
      index = "state",
      model = "within"
    ) %>% 
    summary()
}

names(models_by_social_capital_measure_one_way) <- soc_cap_measures

```

```{r}
models_by_social_capital_measure_one_way
```

# Index and sub-index quarterly regressions complete two-way

```{r}

models_by_social_capital_measure_q1 <- list()

for (i in seq_along(soc_cap_measures)) {
  # for(i in 1:1)  {
  # sc_i <- soc_cap_measures[i]
  
  models_by_social_capital_measure_q1[[i]] <-
    df %>%
    filter(date >= "2020-03-01", date < "2020-06-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2),
      data = .,
      effect = "twoways",
      index = "state",
      model = "within"
    ) %>%
    summary()
}

names(models_by_social_capital_measure_q1) <- soc_cap_measures

```

```{r}
models_by_social_capital_measure_q2 <- list()

for (i in seq_along(soc_cap_measures)) {
  # for(i in 1:1)  {
  # sc_i <- soc_cap_measures[i]
  
  models_by_social_capital_measure_q2[[i]] <-
    df %>%
    filter(date >= "2020-06-01", date < "2020-09-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2),
      data = .,
      effect = "twoways",
      index = "state",
      model = "within"
    ) %>%
    summary()
}

names(models_by_social_capital_measure_q2) <- soc_cap_measures
```

```{r}
models_by_social_capital_measure_q3 <- list()

for (i in seq_along(soc_cap_measures)) {
  # for(i in 1:1)  {
  # sc_i <- soc_cap_measures[i]
  
  models_by_social_capital_measure_q3[[i]] <-
    df %>%
    filter(date >= "2020-09-01", date < "2020-12-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2),
      data = .,
      effect = "twoways",
      index = "state",
      model = "within"
    ) %>%
    summary()
}

names(models_by_social_capital_measure_q3) <- soc_cap_measures
```

```{r}
models_by_social_capital_measure_q4 <- list()

for (i in seq_along(soc_cap_measures)) {
  # for(i in 1:1)  {
  # sc_i <- soc_cap_measures[i]
  
  models_by_social_capital_measure_q4[[i]] <-
    df %>%
    filter(date >= "2020-12-01", date < "2021-03-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2),
      data = .,
      effect = "twoways",
      index = "state",
      model = "within"
    ) %>%
    summary()
}

names(models_by_social_capital_measure_q4) <- soc_cap_measures
```

# \# Index and sub-index quarterly regressions complete one-way

```{r}
models_by_social_capital_measure_one_way_q1 <- list()

for (i in seq_along(soc_cap_measures)) {
  models_by_social_capital_measure_one_way_q1[[i]] <-
    df %>%
    filter(date >= "2020-03-01", date < "2020-06-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2) +
        median_household_income +
        percent_adults_with_ba +
        percent_in_fair_or_poor_health +
        percent_black +
        density +
        percent_rural +
        percent_65 +
        average_temperature +
        governor_political_affiliation,
      data = .,
      effect = "time",
      index = "state",
      model = "within"
    ) %>% 
    summary()
}

names(models_by_social_capital_measure_one_way_q1) <- soc_cap_measures

models_by_social_capital_measure_one_way_q2 <- list()

for (i in seq_along(soc_cap_measures)) {
  models_by_social_capital_measure_one_way_q2[[i]] <-
    df %>%
    filter(date >= "2020-06-01", date < "2020-09-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2) +
        median_household_income +
        percent_adults_with_ba +
        percent_in_fair_or_poor_health +
        percent_black +
        density +
        percent_rural +
        percent_65 +
        average_temperature +
        governor_political_affiliation,
      data = .,
      effect = "time",
      index = "state",
      model = "within"
    ) %>% 
    summary()
}

names(models_by_social_capital_measure_one_way_q2) <- soc_cap_measures

models_by_social_capital_measure_one_way_q3 <- list()

for (i in seq_along(soc_cap_measures)) {
  models_by_social_capital_measure_one_way_q3[[i]] <-
    df %>%
    filter(date >= "2020-09-01", date < "2020-12-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2) +
        median_household_income +
        percent_adults_with_ba +
        percent_in_fair_or_poor_health +
        percent_black +
        density +
        percent_rural +
        percent_65 +
        average_temperature +
        governor_political_affiliation,
      data = .,
      effect = "time",
      index = "state",
      model = "within"
    ) %>% 
    summary()
}

names(models_by_social_capital_measure_one_way_q3) <- soc_cap_measures


models_by_social_capital_measure_one_way_q4 <- list()

for (i in seq_along(soc_cap_measures)) {
  models_by_social_capital_measure_one_way_q4[[i]] <-
    df %>%
    filter(date >= "2020-12-01", date < "2021-03-01") %>%
    plm(
      cases_new ~ stringency_index_lag * get(soc_cap_measures[i]) +
        I(stringency_index_lag ^ 2) +
        median_household_income +
        percent_adults_with_ba +
        percent_in_fair_or_poor_health +
        percent_black +
        density +
        percent_rural +
        percent_65 +
        average_temperature +
        governor_political_affiliation,
      data = .,
      effect = "time",
      index = "state",
      model = "within"
    ) %>% 
    summary()
}

names(models_by_social_capital_measure_one_way_q4) <- soc_cap_measures

```

# Index and sub-index quarterly regressions complete two-way

```{r eval=FALSE}

# can't make nested loops work

q_dates <- tribble(
  ~begin, ~end,
  "2020-03-01", "2020-06-01",
  "2020-06-01", "2020-09-01",
  "2020-09-01", "2020-12-01",
  "2020-12-01", "2021-03-01"
)

models_by_quarter <- list()
models_by_social_capital <- list()

for(i in 1:nrow(q_dates)) {
  models_by_quarter[[length(models_by_quarter) + 1]] <- 
  for (j in seq_along(soc_cap_measures)) {
    c(models_by_social_capital[[j]] <- 
      df %>%
      filter(date >= q_dates[[1]][i], date < q_dates[[2]][i]) %>%
      plm(
        cases_new ~ stringency_index_lag * get(soc_cap_measures[j]) +
          I(stringency_index_lag ^ 2),
        data = .,
        effect = "twoways",
        index = "state",
        model = "within"
      ) %>%
      summary()
    )
    
  }
  
  
}


```

```{r eval=FALSE}
models_by_social_capital[1]
models_by_quarter[1]
```

```{r eval=FALSE}

df %>%
  mutate(quarter =
           as_factor(
             case_when(
               date >= "2020-03-10" & date < "2020-03-01" ~ "all",
               date >= "2020-03-10" & date < "2020-06-01" ~ "q1",
               date >= "2020-06-01" & date < "2020-09-01" ~ "q2",
               date >= "2020-09-01" & date < "2020-12-01" ~ "q3",
               date >= "2020-12-01" & date < "2021-03-01" ~ "q4"
             )
           ))
  
  
  

```

```{r eval=FALSE}
df_by_index_quarter <-
  df %>%
  mutate(quarter =
           as_factor(
             case_when(
               date >= "2020-03-10" & date < "2020-06-01" ~ "q1",
               date >= "2020-06-01" & date < "2020-09-01" ~ "q2",
               date >= "2020-09-01" & date < "2020-12-01" ~ "q3",
               date >= "2020-12-01" & date < "2021-03-01" ~ "q4"
               )
           )) %>%
  select(-contains("dummy")) %>% 
  pivot_longer(soc_cap_measures,
               names_to = "social_k_name",
               values_to = "social_k_index") %>%
  group_by(quarter) %>%
  nest() 
  
```

```{r eval=FALSE}

df_by_index_quarter 

df_by_index_quarter$data[1] %>% 
  glimpse()
```

# Two-way models 

## all observations

Coefficients for two-way models. In order stringency_index_lag, stringency_index_lag^2^ , stringency_index_lag \* social_capital_index

```{r}
models_by_social_capital_measure %>% 
  map("coefficients") %>% 
  map(`[`, 1:3) 
```

## by quarter

### first quarter March 2020 - May 2020

```{r}
models_by_social_capital_measure_one_way_q1 %>% 
  map("coefficients") %>% 
  map(`[`, 1:3)
```

### second quarter June 2020 - Aug 2020

```{r}
models_by_social_capital_measure_one_way_q2 %>% 
  map("coefficients") %>% 
  map(`[`, 1:3)
```

### third quarter Sep 2020 - Nov 2020

```{r}
models_by_social_capital_measure_one_way_q3 %>% 
  map("coefficients") %>% 
  map(`[`, 1:3)
```

### 4th quarter

```{r}
models_by_social_capital_measure_one_way_q4 %>% 
  map("coefficients") %>% 
  map(`[`, 1:3)
```

# One-way models

## all observations

Coefficients for 
  - stringency_index_lag
  - social_capital_measure
  - median_household_income 
  - percent_adults_with_ba 
  - percent_in_fair_or_poor_health 
  - percent_black 
  - density 
  - percent_rural 
  - percent_65 
  - average_temperature 
  - governor_political_affiliation
  

```{r}
models_by_social_capital_measure_one_way %>% 
  map(coefficients) %>% 
  map(`[`, 1:13)
```

## by quarter

### Mar 2020 - May 2020

```{r}
models_by_social_capital_measure_one_way_q1 %>% 
  map(coefficients) %>% 
  map(`[`, 1:13)
```

### Jun 2020 - Aug 2020

```{r}
models_by_social_capital_measure_one_way_q2 %>% 
  map(coefficients) %>% 
  map(`[`, 1:13)
```

### Sep 2020 - Nov 2020

```{r}
models_by_social_capital_measure_one_way_q3 %>% 
  map(coefficients) %>% 
  map(`[`, 1:13)
```

### Dec 2020 - Mar 2021

```{r}
models_by_social_capital_measure_one_way_q4 %>% 
  map(coefficients) %>% 
  map(`[`, 1:13)
```
